---
title: "MBissonnette Phase 1 Project"
author: "me"
date: "2023-02-22"
output:
  word_document: default
  html_document: default
---
Needed libraries  
```{r}
library(tidyverse)
library(tidymodels)
library(dplyr)
library(GGally) #ggcorr and ggpairs
library(ggcorrplot) #correlation plot alternative
library(glmnet) #for Lasso, ridge, and elastic net models 
library(gridExtra) #create grids of plots
library(car) #for the VIF function
library(MASS) #access to forward and backward selection algorithms
library(leaps) #best subset selection
library(lmtest) 
library(splines) #for nonlinear fitting
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(naniar) #visualizing missingness
library(skimr) #alternative way to view dataset summaries
library(UpSetR) #visualizing missingness
```


Read-in the dataset
```{R}
amesstudent = read_csv("ames_student-1.csv",show_col_types = FALSE)
```

Structure and summary. 
```{r}
str(amesstudent)
summary(amesstudent)
glimpse(amesstudent)
skim(amesstudent)
```
Select some important variables.   
```{r}
amesstudent2 = amesstudent %>% dplyr::select("Neighborhood", "Bldg_Type", "House_Style", "Overall_Qual",  "Year_Built", "Year_Remod_Add", "Total_Bsmt_SF", "Full_Bath", "Bedroom_AbvGr", "Fireplaces", "Garage_Type", "Garage_Cars", "Above_Median")
```

Structure and summary. 
```{r}
str(amesstudent2)
summary(amesstudent2)
glimpse(amesstudent2)
skim(amesstudent2)
```

```{r}
amesstudent2 = amesstudent2 %>% mutate(across(where(is.character),as_factor))
```
Structure and summary. 
```{r}
str(amesstudent2)
summary(amesstudent2)
glimpse(amesstudent2)
skim(amesstudent2)
```

Simple view of missingess. As you can see, no missing data.
```{r}
gg_miss_var(amesstudent2)
```

### Data Exploration
Begin exploring the data by looking at a plot of our response variable only (choose a histogram for a single quantitative variable)  

```{r}
ggplot(amesstudent2, aes(x=Above_Median)) + geom_bar() + theme_bw()
```
Next we look at correlation. This is a logical step since almost all of our variables are quantitative.Year_Built has highest correlation

Our Y (response) variable in this dataset is "Above_Median".  Let's look at ggpairs plot for visualization and correlation. There is a strong correlation between Year_Remod_Add and Above_Median 
```{r}
ggpairs(amesstudent2,cardinality_threshold = 28,upper = list(continuous=wrap("cor", size=2.0)),lower = list(combo="count"))
```
Alternatively:  Year_Built, Year_Remod_Add, and Total_Bmst_SF all appear to impact Above_Median.
```{r}
p1 = ggplot(amesstudent2, aes(x=Neighborhood,y=Above_Median)) + geom_point(alpha = 0.1) #changing alpha is helpful when many points may overlap
p2 = ggplot(amesstudent2, aes(x=Bldg_Type,y=Above_Median)) + geom_boxplot()
p3 = ggplot(amesstudent2, aes(x=House_Style,y=Above_Median)) + geom_boxplot()
p4 = ggplot(amesstudent2, aes(x=Overall_Qual,y=Above_Median)) + geom_boxplot()
p5 = ggplot(amesstudent2, aes(x=Year_Built,y=Above_Median)) + geom_boxplot()
p6 = ggplot(amesstudent2, aes(x=Year_Remod_Add,y=Above_Median)) + geom_boxplot()
p7 = ggplot(amesstudent2, aes(x=Total_Bsmt_SF,y=Above_Median)) + geom_boxplot()
p8 = ggplot(amesstudent2, aes(x=Full_Bath,y=Above_Median)) + geom_boxplot()
p9 = ggplot(amesstudent2, aes(x=Bedroom_AbvGr,y=Above_Median)) + geom_boxplot()
p10 = ggplot(amesstudent2, aes(x=Fireplaces,y=Above_Median)) + geom_boxplot()
p11 = ggplot(amesstudent2, aes(x=Garage_Type,y=Above_Median)) + geom_boxplot()
p12 = ggplot(amesstudent2, aes(x=Garage_Cars,y=Above_Median)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,ncol=2)
```
##Create a table to look at correlation between Neighborhood and Above_Median. Can see there is a correlation between value of home and neighborhood.
```{r}
amesstudent3<- table(amesstudent2$Neighborhood, amesstudent2$Above_Median)
  amesstudent3
```

Closer look at "Year_Built" and "Above_Median".  This shows that the newer the home, the above_median is higher.
```{r}
ggplot(amesstudent2, aes(x=Year_Built,y=Above_Median)) + geom_boxplot() + theme_bw()
ggplot(amesstudent2, aes(x=Year_Built,y=Above_Median)) + geom_boxplot() + 
  geom_jitter() + theme_bw()
```

Closer look at "Year_Remod_Add" and "Above_Median".  This shows that the newer the remodel, the above_median is higher.
```{r}
ggplot(amesstudent2, aes(x=Year_Remod_Add,y=Above_Median)) + geom_boxplot() + theme_bw()
ggplot(amesstudent2, aes(x=Year_Remod_Add,y=Above_Median)) + geom_boxplot() + 
  geom_jitter() + theme_bw()
```

Closer look at "Total_Bsmt_SF" and "Above_Median".  This shows that the more basement square footage, the above_median is higher.
```{r}
ggplot(amesstudent2, aes(x=Total_Bsmt_SF,y=Above_Median)) + geom_boxplot() + theme_bw()
ggplot(amesstudent2, aes(x=Total_Bsmt_SF,y=Above_Median)) + geom_boxplot() + 
  geom_jitter() + theme_bw()
```

##As seen above, Above_Median is highly correlated to Neighborhood, Year_Built, Year_Remod_Add, and Total_Bsmt_SF. 